Accelerating Burn Speed of Birch Wood Popsicle Sticks

### 2022-05-01 paper/paper.Rmd
Update spelling

If we soak popsicle sticks in ethanol alcohol, isopropyl alcohol, and acetone and burn them, comparing distance burned against a dry popsicle stick, then we think that using any liquid accelerant would be faster than no liquid accelerant. We also think that in comparing individual treatments, there would be no significant difference between the accelerants, but there would be a significant difference between no accelerant and with accelerant.

@@

For the experiment, we used birch wood popsicle sticks. We also prepared our treatments, which were the ethanol alcohol, isopropyl alcohol, acetone, and no accelerant. To hold the burning sticks off the ground (so they could be lit from below, since flames rise), we used four mason jars, one for each stick within a block. To light the sticks, we used a barbeque lighter, and for timing, we used an Apple iPhone. For safety purposes, we also had extra water.

@@

Data were cleaned using Python version 3.7 (Python, Wilmington, DE) and Pandas version 1.0.3. Data were analyzed using R version 4.0.3 (The R Project for Statistical Computing, Indianapolis, IN). The Skillings-Mack procedure was run in R using the NSM3 package version 1.16. The Wilcoxon rank-sum tests were run using the stats package in R.

@@

In Figure 4, there are flaired tails that indicate an overdispersed data distribution compared to a normal distribution. That is, the distribution was "squished" toward the center compared to the normal distribution.

@@

$$
H_0: [\alpha_1 = \alpha_2 = \alpha_3 = \alpha_4],
$$

$$
H_A: [\alpha_1, \alpha_2, \alpha_3, \alpha_4 \text{ not all equal}],
$$

@@

| Comparison        | Actual | Medians of Simulated U’s | $\frac{0.1}{6}$ Percentile | $1 - \frac{0.1}{6}$ Percentile |
|-------------------|--------|--------------------------|----------------|-----------------|
| Ethanol-Dry       | 2.5    | 18.5                     | 4.833333       | 30.66667        |
| Ethanol-Acetone   | 2.0    | 18.0                     | 5.000000       | 30.50000        |
| Ethanol-Isopropyl | 35.0   | 18.0                     | 5.500000       | 30.50000        |
| Acetone-Dry       | 17.0   | 17.5                     | 4.333333       | 31.16667        |
| Acetone-Isopropyl | 19.0   | 18.0                     | 4.333333       | 31.16667        |
| Isopropyl-Dry     | 23.0   | 18.0                     | 4.000000       | 30.00000        |

@@

In Table 1, we found that the simulated U's are not statistically different from the actual U's for the acetone-dry, acetone-isopropyl, and isopropyl-dry combinations. This means that there is no statistically significant difference between the results for isopropyl alcohol, acetone, and no accelerant.

>>>>

If we soak popsicle sticks in acetone, ethanol alcohol, and isopropyl alcohol and burn them, comparing distance burned against a dry popsicle stick, then we think that using any liquid accelerant would be faster than no liquid accelerant. We also think that in comparing individual treatments, there would be no significant difference between the accelerants, but there would be a significant difference between no accelerant and with accelerant.

@@

For the experiment, we used birch wood popsicle sticks. We also prepared our treatments, which were the acetone, ethanol alcohol, isopropyl alcohol, and no accelerant. To hold the burning sticks off the ground (so they could be lit from below, since flames rise), we used four mason jars, one for each stick within a block. To light the sticks, we used a barbecue lighter, and for timing, we used an Apple iPhone. For safety purposes, we also had extra water.

@@

Data were cleaned using Python version 3.7 (Python, Wilmington, DE) and Pandas version 1.0.3. Data were analyzed using R version 4.0.3 (The R Project for Statistical Computing, Indianapolis, IN). The Skillings-Mack procedure was run in R using the NSM3 package, version 1.16. The Wilcoxon rank-sum tests were run using the stats package in R.

@@

In Figure 4, there are flared tails that indicate an overdispersed data distribution compared to a normal distribution. That is, the distribution was "squished" toward the center compared to the normal distribution.

@@

$$
H_0: [\alpha_\text{acetone} = \alpha_\text{dry} = \alpha_\text{ethanol} = \alpha_\text{isopropyl}],
$$

$$
H_A: [\alpha_\text{acetone}, \alpha_\text{dry}, \alpha_\text{ethanol}, \alpha_\text{isopropyl} \text{ not all equal}],
$$

@@

| Comparison        | Actual $U$ | Medians of Simulated $U$’s | $\frac{0.1}{6}$ Percentile | $1 - \frac{0.1}{6}$ Percentile |
|-------------------|--------|--------------------------|----------------|-----------------|
| Ethanol-Dry       | 2.5    | 18.5                     | 4.833333       | 30.66667        |
| Ethanol-Acetone   | 2.0    | 18.0                     | 5.000000       | 30.50000        |
| Ethanol-Isopropyl | 35.0   | 18.0                     | 5.500000       | 30.50000        |
| Acetone-Dry       | 17.0   | 17.5                     | 4.333333       | 31.16667        |
| Acetone-Isopropyl | 19.0   | 18.0                     | 4.333333       | 31.16667        |
| Isopropyl-Dry     | 23.0   | 18.0                     | 4.000000       | 30.00000        |

@@

In Table 1, we found that the simulated $U$'s are not statistically different from the actual $U$'s for the acetone-dry, acetone-isopropyl, and isopropyl-dry combinations. This means that there is no statistically significant difference between the results for isopropyl alcohol, acetone, and no accelerant.
